Forecasting the power of Higher Order Weak Lensing Statistics with automatically differentiable simulations

### 2023-06-06 paper/journal_paper.tex
Submitted version

\author{Denise Lanzieri \inst{1}
\and
Fran\c{c}ois Lanusse \inst{2}
\and
Chirag Modi \inst{3}
\and 
Benjamin Horowitz \inst{4, 5}
\and
Joachim Harnois-Déraps \inst{6}
\and
Jean-Luc Starck \inst{2}
\and
The LSST Dark Energy Science Collaboration (LSST DESC)
}
\institute{Université Paris Cité, Université Paris-Saclay, CEA, CNRS, AIM, F-91191, Gif-sur-Yvette, France
\and
Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France;
\and
Center for Computational Astrophysics,
Center for Computational Mathematics,
Flatiron Institute, New York, NY 10010, USA;
\and
Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, 94720, CA, USA;
\and
Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544, USA
\and
School of Mathematics, Statistics and Physics, Newcastle University, Herschel Building, NE1 7RU Newcastle-upon-Tyne, UK
}

@@

   \keywords{methods: statistical -- dark energy
               }

@@

\section{Weak Lensing Modeling}\label{weak_lesning_modeling}
\subsection{{Cosmic shear}}
Weak gravitational lensing is a powerful probe to infer the projected matter density distribution between an observer and a source.
The effect of gravitational lensing 
can be quantified in term of the separation vector \textbf{x} between two light rays separated by an angle $\boldsymbol{\theta}$:
\begin{align}\label{sep}
   & \textbf{x}(\boldsymbol{\theta},\chi) =
    f_k(\chi)\boldsymbol{\theta} + \\
   - \frac{2}{c^2}
   \int_0^{\chi}
    &  \text{d}\chi'
    f_k(\chi-\chi')
    [ \boldsymbol{\nabla_{\bot}}\Phi( \textbf{x}(\boldsymbol{\theta},\chi'),\chi')-
    \boldsymbol{\nabla_{\bot}}\Phi^{(0)}(\chi')
    ],
\end{align}
with $\Phi$ and $\Phi^{0}$ the gravitational potential along the two light rays, $f_k(\chi)$ and $\chi$ the angular and radial comoving distance.

@@

DESC acknowledges ongoing support from the IN2P3 (France), the STFC 
(United Kingdom), and the DOE, NSF, and LSST Corporation (United States).  
DESC uses resources of the IN2P3 Computing Center 
(CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la
Recherche Scientifique; the National Energy Research Scientific Computing
Center, a DOE Office of Science User Facility supported under Contract 
No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BEIS National 
E-infrastructure capital grants; and the UK particle physics grid, supported
by the GridPP Collaboration.  This work was performed in part under DOE 
Contract DE-AC02-76SF00515.
This work was granted access to the HPC/AI resources of IDRIS under the allocation 2022-AD011013922 made by GENCI.
BH is supported by the AI Accelerator program of the Schmidt Futures Foundation; JHD acknowledges support from an STFC Ernest Rutherford Fellowship (project reference ST/S004858/1); The presented work used computing resources provided by DESC at the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DEAC02-05CH11231; 
This paper has undergone internal review in the LSST Dark Energy Science Collaboration. We thank the internal reviewers, Virginia Ajani, William Coulton, and Jia Liu for their constructive comments.
Author contributions to this work are as follows: Denise Lanzieri led the analysis and the writing of the paper.  Jean-Luc Starck provided his expertise on wavelet peak counting and higher-order weak lensing statistics. All the authors provided feedback on
the text.
\end{acknowledgements}

>>>>

\author{Denise Lanzieri \inst{1}\fnmsep\thanks{ Contact: denise.lanzieri@cea.fr}
\and
Fran\c{c}ois Lanusse \inst{2}
\and
Chirag Modi \inst{3}
\and 
Benjamin Horowitz \inst{4, 5}
\and
Joachim Harnois-Déraps \inst{6}
\and
Jean-Luc Starck \inst{2}
\and
The LSST Dark Energy Science Collaboration (LSST DESC)
}
\institute{Université Paris Cité, Université Paris-Saclay, CEA, CNRS, AIM, F-91191, Gif-sur-Yvette, France
\and
Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France
\and
Center for Computational Astrophysics,
Center for Computational Mathematics,
Flatiron Institute, New York, NY 10010, USA
\and
Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, 94720, CA, USA
\and
Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544, USA
\and
School of Mathematics, Statistics and Physics, Newcastle University, Herschel Building, NE1 7RU Newcastle-upon-Tyne, UK
}

@@

\keywords{methods: statistical -- cosmology: large-scale structure of Universe --gravitational lensing:weak}

@@

\section{Weak Lensing Modeling}\label{weak_lesning_modeling}
\subsection{{Cosmic shear}}
Weak gravitational lensing is a powerful probe to infer the distribution of matter density between an observer and a source.
The effect of gravitational lensing 
can be quantified in term of the separation vector \textbf{x} between two light rays separated by an angle $\boldsymbol{\theta}$:
\begin{align}\label{sep}
   & \textbf{x}(\boldsymbol{\theta},\chi) =
    f_k(\chi)\boldsymbol{\theta} + \\
   - \frac{2}{c^2}
   \int_0^{\chi}
    &  \text{d}\chi'
    f_k(\chi-\chi')
    [ \boldsymbol{\nabla_{\bot}}\Phi( \textbf{x}(\boldsymbol{\theta},\chi'),\chi')-
    \boldsymbol{\nabla_{\bot}}\Phi^{(0)}(\chi')
    ],
\end{align}
with $\Phi$ and $\Phi^{0}$ the gravitational potential along the two light rays, $f_k(\chi)$ and $\chi$ the angular and radial comoving distance.

@@

DESC acknowledges ongoing support from the IN2P3 (France), the STFC 
(United Kingdom), and the DOE, NSF, and LSST Corporation (United States).  
DESC uses resources of the IN2P3 Computing Center 
(CC-IN2P3--Lyon/Villeurbanne - France) funded by the Centre National de la
Recherche Scientifique; the National Energy Research Scientific Computing
Center, a DOE Office of Science User Facility supported under Contract 
No.\ DE-AC02-05CH11231; STFC DiRAC HPC Facilities, funded by UK BEIS National 
E-infrastructure capital grants; and the UK particle physics grid, supported
by the GridPP Collaboration.  This work was performed in part under DOE 
Contract DE-AC02-76SF00515.
This work was granted access to the HPC/AI resources of IDRIS under the allocation 2022-AD011013922 made by GENCI.
BH is supported by the AI Accelerator program of the Schmidt Futures Foundation; JHD acknowledges support from an STFC Ernest Rutherford Fellowship (project reference ST/S004858/1); The presented work used computing resources provided by DESC at the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DEAC02-05CH11231; 
This paper has undergone internal review in the LSST Dark Energy Science Collaboration. We thank the internal reviewers, Virginia Ajani, William Coulton, and Jia Liu for their constructive comments.
Author contributions to this work are as follows: Denise Lanzieri led the analysis and the writing of the paper. Fran\c{c}ois Lanusse designed the project, contributed to code and text development.
Chirag Modi contributed to the PM code and text development.
Benjamin Horowitz contributed to the text and made minor code developments. Joachim Harnois-Déraps contributed to the intrinsic alignment modeling, provided general advice on simulations, and contributed to the text.
Jean-Luc Starck provided expertise on wavelet peak counting and higher-order weak lensing statistics. 
\end{acknowledgements}
